---
title: "R Notebook"
output: html_notebook
---

---
title: "R Notebook"
output: html_notebook
---

## Local Database: MySQL on AWS
```{r}
# 1. Library (must be installed prior to loading
library(RMySQL)     ### MySQL
library(RMariaDB)   ### MariaDB
library(DBI)

# 2. Settings
db_user <- 'mu_bai'
db_password <- 'Mc5539899!'
db_name <- 'mu_bai_aws_db'
db_host <- 'cs5200-dbs.c6iyvvr9xuen.us-east-2.rds.amazonaws.com'
db_port <- 3306


# 3. Connect to DB
my_awsDB <-  dbConnect(MySQL(), user = db_user, password = db_password,
                 dbname = db_name, host = db_host, port = db_port)

```


```{r}

dbListTables(my_awsDB)
```

## TASK 1. CREATE TABLES

##create airlines table

```{sql connection=my_awsDB}
DROP TABLE IF EXISTS airlines
```

```{sql connection=my_awsDB}
CREATE TABLE airlines (
	aid NUMERIC NOT NULL,
  code CHAR,
  airline CHAR UNIQUE,
  PRIMARY KEY (aid, airline)
  
)

```

##create airports table

```{sql connection=my_awsDB}
DROP TABLE IF EXISTS airports
```

```{sql connection=my_awsDB}
CREATE TABLE airports (
	pid NUMERIC NOT NULL,
  code CHAR,
  name CHAR UNIQUE,
  city CHAR,
  state CHAR,
  country CHAR,
  PRIMARY KEY (pid, name)
)
```

##Create incidents table

```{sql connection=my_awsDB}
DROP TABLE IF EXISTS incidents
```


```{sql connection=my_awsDB}
CREATE TABLE incidents (
	iid NUMERIC NOT NULL,
  flight_date DATE,
  depPort CHAR,
  arrPort CHAR,
  airline CHAR,
  aircraft CHAR,
  flightPhase CHAR CONSTRAINT chk_flightPhase CHECK (flightPhase IN ('takeoff', 'landing', 'inflight', 'unknown')),
  impact CHAR,
  PRIMARY KEY (iid)
)
```

##Set a foreign key constraint: FOREIGN KEY (depPort) REFERENCES airports (name)
```{sql connection=my_awsDB}
ALTER TABLE incidents ADD CONSTRAINT FOREIGN KEY (airline) REFERENCES airlines(airline);
```
##Set another FOREIGN KEY (airline) REFERENCES airlines (airline)
```{sql connection=my_awsDB}
ALTER TABLE incidents ADD CONSTRAINT FOREIGN KEY (depPort) REFERENCES airports(name);
```

## TASK 2. POPULATING TABLES WITH CSV DATA

## STEP ONE: IMPORT THE CSV FILE BY USING sqldf()
```{r}
BIncident_df <- read.csv("BirdStrikesData.csv")
BIncident_df <- na.omit(BIncident_df)

BIncident_df$aid <- seq(from = 1, to = nrow(BIncident_df)) + 100
BIncident_df
```

```{r}
BIncident_df$FlightDate <- substring(BIncident_df$FlightDate, 1, 
                                               nchar(BIncident_df$FlightDate)-5)


BIncident_df$When..Phase.of.flight <- sub("Landing Roll", "landing", BIncident_df$When..Phase.of.flight)
BIncident_df$When..Phase.of.flight <- sub("Take-off run", "takeoff", BIncident_df$When..Phase.of.flight)
BIncident_df$When..Phase.of.flight <- sub("Climb", "inflight", 
                                                 BIncident_df$When..Phase.of.flight)
BIncident_df$When..Phase.of.flight <- sub("Descent", "inflight", 
                                                 BIncident_df$When..Phase.of.flight)
BIncident_df$When..Phase.of.flight <- sub("Approach", "inflight", 
                                                 BIncident_df$When..Phase.of.flight)
BIncident_df$When..Phase.of.flight <-  sub("Parked", "inflight", 
                                                 BIncident_df$When..Phase.of.flight)
BIncident_df$When..Phase.of.flight <-  sub("Taxi", "unknown", 
                                                 BIncident_df$When..Phase.of.flight)
BIncident_df$When..Phase.of.flight <-  sub("^$", "unknown", 
                                                 BIncident_df$When..Phase.of.flight)
BIncident_df
```


##Create dataframe for incidents table
```{r}
df.incidents <- data.frame(iid = BIncident_df$aid,
                         flight_date = BIncident_df$FlightDate,
                         depPort = BIncident_df$Airport..Name,
                         arrPort = BIncident_df$Airport..Name,
                         airline = BIncident_df$Aircraft..Airline.Operator,
                         aircraft = BIncident_df$Aircraft..Make.Model,
                         flightPhase = BIncident_df$When..Phase.of.flight,
                         impact = BIncident_df$Effect..Indicated.Damage)
df.incidents
```
##Create dataframe for airlines table
```{r}
library(sqldf)
options(sqldf.driver = "SQLite")
df.airlines <- sqldf::sqldf("SELECT 1 as aid, '' as code, 
`Aircraft..Airline.Operator` as airlines from `BIncident_df` 
WHERE airlines NOT LIKE '' group by airlines ")

num_airines <- nrow(df.airlines)
df.airlines[,1] <- seq(1, num_airines)
df.airlines <- na.omit(df.airlines)
df.airlines 
```
##Create dataframe for airports table
```{r}
df.airports <- sqldf::sqldf("SELECT 1 as pid, '' as code, 
`Airport..Name` as name, '' as city, `Origin.State` as state, 'USA' as country from `BIncident_df`
WHERE name NOT LIKE '' group by name")

num_airports <- nrow(df.airports)
df.airports[,1] <- seq(1, num_airports)
df.airports <- na.omit(df.airports)
df.airports
```

## BULK LOADING DATA TO EACH TABLE

```{sql connection=my_awsDB}
DELETE FROM incidents;
```

```{sql connection=my_awsDB}
DELETE FROM airlines;
```

```{sql connection=my_awsDB}
DELETE FROM airports;
```


```{r}
dbWriteTable(my_awsDB, "incidents", df.incidents, overwrite = T, row.names =F)
```

```{r}
dbWriteTable(my_awsDB, "airlines", df.airlines, overwrite = T, row.names =F)
```

```{r}
dbWriteTable(my_awsDB, "airports", df.airports, overwrite = T, row.names =F)
```


##Question 3
##Check wether incidents table is populated in my_awsDB
```{sql connection=my_awsDB}
select * from `incidents` limit 10;
```

##Check wether airlines table is populated in my_awsDB
```{sql connection=my_awsDB}
select * from `airlines` limit 10;
```
##Check wether airports table is populated in my_awsDB
```{sql connection=my_awsDB}
select * from `airports` limit 10;
```
## Question 4

```{sql connection=my_awsDB}
SELECT num_of_strikes, airline, arrPort, flightPhase from
(SELECT count(iid) as num_of_strikes, airline, arrPort, flightPhase from incidents group by airline) AS A
WHERE arrPort = "LAGUARDIA NY" group by airline;

```


## Question 5

```{sql connection=my_awsDB}
SELECT num_of_strikes as Top_Strikes, Airport, flightPhase from (SELECT COUNT(iid) as num_of_strikes, 
depPort as Airport, flightPhase from incidents WHERE airline NOT LIKE
"%BUSINESS%" or "%PRIVATE%"or "%MILITARY%" group by Airport) AS A
ORDER BY Top_Strikes desc limit 1;
```

## Question 6
```{sql connection=my_awsDB}
SELECT Year, Total_Number_of_Strikes, Airport, Flights 
FROM (SELECT depPort as Airport, airline as Flights, COUNT(iid) 
AS Total_Number_of_Strikes, SUBSTRING(flight_date, length(flight_date)-3) AS Year
FROM incidents group by Year) AS A
WHERE Year = '2005' or Year = '2006' or Year = '2007' or Year = '2008'
or Year = '2009' or Year = '2010' or Year = '2011';
```



##Question 7

```{r}
df <- dbReadTable(my_awsDB, "incidents")
```

```{r}
library(sqldf)
options(sqldf.driver = "SQLite")
df <- sqldf("SELECT Year, Total_Number_of_Strikes, Airport, Flights 
FROM (SELECT depPort as Airport, airline as Flights, COUNT(iid) 
AS Total_Number_of_Strikes, SUBSTRING(flight_date, length(flight_date)-3) AS Year
FROM df group by Year) AS A
WHERE Year = '2005' or Year = '2006' or Year = '2007' or Year = '2008'
or Year = '2009' or Year = '2010' or Year = '2011'")
df
```


```{r}
library(ggplot2)
line_chart <- ggplot(df, aes(x=`Year`, y=`Total_Number_of_Strikes`, group = 1)) +
geom_line() +
labs(x = "Year", y = "Number of strikes") +
ggtitle("Number of bird strikes happened by year") +
geom_text(aes(label=`Total_Number_of_Strikes`, vjust= -0.5))
line_chart
```

## Question 8 
Created a stored procedure called DeleteIncidentByAirport with a given name of
an airport same as depPort to delete rows that have matching quality.

```{sql connection=my_awsDB}
DROP PROCEDURE IF EXISTS DeleteIncidentByAirport;
```

```{sql connection=my_awsDB}
CREATE PROCEDURE DeleteIncidentByAirport(
 IN AirportName CHAR(255)
)
BEGIN 
  DELETE FROM `incidents` 
  WHERE depPort = AirportName;
END
```


##Invoke the stored procedure named DeleteIncidentByAirport by using CALL
```{sql connection=my_awsDB}
CALL DeleteIncidentByAirport("LAGUARDIA NY")
```

##Test case result: first row and so on with airport name "LAGUARDIA NY" are
##deleted as expected as iid 101 is no longer in the table incidents.
```{sql connection=my_awsDB}
SELECT * FROM `incidents`;
```


```{r}
dbDisconnect(my_awsDB)
```


