---
title: "HM3"
author: "Mu Cheng"
date: "3/1/2021"
output: pdf_document
---

## Problem 1
```{r echo=TRUE, message=FALSE, warning=FALSE}
getwd()
setwd("~/Desktop/5110/HM3")
library("readr")
library(plyr)
library(dplyr)

mydir = "epa-aqi-data-annual"
myfiles <- list.files(path=mydir, pattern="*.csv", full.names=TRUE) %>%
  lapply(read_csv) %>%
  bind_rows()

```
## 1st visulization with boxplot
```{r echo=TRUE, message=FALSE, warning=FALSE}
library(ggplot2)
ggplot(myfiles, mapping=aes(x=as.factor(Year), y=`Median AQI`, color=`Year`)) + 
  geom_boxplot() + labs(x="Years", 
                        y="Median AQI", 
                        title=
"Median AQI decreased incrementally with fewer outliers on high during 1980-2019 ") + theme_minimal() + 
  theme(axis.text.x = element_text(angle=90, vjust=0.5,  hjust=1))

```
## 2nd Visulization with Scatter-plot. 
The second graph is made to assure the conclusion addressed on outliers that 
are getting fewer than before is true in the above analysis.

```{r echo=TRUE, message=FALSE, warning=FALSE}
library(ggplot2)
ggplot(myfiles, mapping=aes(x=as.factor(Year), y=`Median AQI`, color=`Year`)) + 
  geom_point() + labs(x="Years", 
                      y="Median AQI", 
                      title=
"Median AQI decreases incrementally with fewer outliers on high during 1980-2019") + theme_minimal() + 
  theme(axis.text.x = element_text(angle=90, vjust=0.5,  hjust=1))

```
## Problem 2
```{r echo=TRUE, message=FALSE, warning=FALSE}

lowerLetters <- function(s) {
    paste(tolower(substring(s, 1, 20)))
}

myfiles$State <- lowerLetters(myfiles$State)


decade1 <- filter(myfiles, `Year` >= 1980 & `Year` <= 1989)

decade2 <- filter(myfiles, `Year` >= 1990 & `Year` <= 1999)

decade3 <- filter(myfiles, `Year` >= 2000 & `Year` <= 2009)

decade4 <- filter(myfiles, `Year` >= 2010 & `Year` <= 2019)
```

```{r echo=TRUE, message=FALSE, warning=FALSE}
decade1_state <-select(decade1, `State`, `Median AQI`) %>% 
  group_by(`State`) %>% 
  summarize_each(funs(mean(`Median AQI`, na.rm=TRUE)))

decade2_state <-select(decade2, `State`, `Median AQI`) %>% 
group_by(`State`) %>% 
  summarize_each(funs(mean(`Median AQI`, na.rm=TRUE)))

decade3_state <-select(decade3, `State`, `Median AQI`) %>% 
  group_by(`State`) %>% 
  summarize_each(funs(mean(`Median AQI`, na.rm=TRUE)))

decade4_state <-select(decade4, `State`, `Median AQI`) %>% group_by(`State`) %>% 
  summarize_each(funs(mean(`Median AQI`, na.rm=TRUE)))
```

```{r echo=TRUE, message=FALSE, warning=FALSE}
library(maps)

us_map <- map_data("state")

us_map1 <-inner_join(us_map, decade1_state, by=c("region"="State"))

ggplot() +
  geom_polygon(data = us_map1, aes(x = long, 
                                   y = lat, group = group, fill=`Median AQI`)) +
  coord_map() + labs(x="Longtitude",
                     y="latitude",
                     title=
"Good Overall AQI with 5 states having lower than 30 AQI during 1980-1989")

```
```{r}
us_map2 <-inner_join(us_map, decade2_state, by=c("region"="State"))

ggplot() +
  geom_polygon(data = us_map2, aes(x = long, 
                                   y = lat, group = group, fill=`Median AQI`)) +
  coord_map() + labs(x="Longtitude",
                     y="latitude", title=
"The highest AQI reduced to 50s but the overall AQI grew higher during 1900-1999")
```
```{r}
us_map3 <-inner_join(us_map, decade3_state, by=c("region"="State"))

ggplot() +
  geom_polygon(data = us_map3, aes(x = long, 
                                   y = lat, group = group, fill=`Median AQI`)) +
  coord_map() + labs(x="Longtitude",
                     y="latitude",
                     title=
"Higher AQI in south and lower AQI in northwest during 2000-2009")
```

```{r}
us_map4 <-inner_join(us_map, decade4_state, by=c("region"="State"))

ggplot() +
  geom_polygon(data = us_map4, aes(x = long, 
                                   y = lat, group = group, fill=`Median AQI`)) +
  coord_map() + labs(x="Longtitude",
                     y="latitude", title=
"Overall quality of air is improved with highest AQI only at 50s during 2010-2019")
```

## Problem 3
```{r echo=TRUE, message=FALSE, warning=FALSE}
getwd()
setwd("~/Desktop/5110/HM3/ddf--gapminder--systema_globalis-master")
library("readr")
library(ggplot2)

country_entities <- read_csv("ddf--entities--geo--country.csv")

country_region_match <- select (country_entities, `name`, `world_4region`)

a <- na.omit(country_region_match)

country_num_each_region <- ggplot(data=a, mapping=aes(x=`world_4region`, fill=`world_4region`)) + 
geom_bar() + labs(x="world 4 region",
                  y="count",
                  title="Number of Countries in Each Region")
country_num_each_region

```

## Problem 4

In Africa:The overall infant mortality rate has decreased during 1950-2015. 
Twice had the rate happened to increase, they were at 1875-1900 and late 
1930s-1960s.

In America:The overall infant mortality rate has decreased during 1950-2015. 
One major increase happened at 1875-1912 time-period.

In Asia:The overall infant mortality rate has decreased during 1950-2015. 
One major increase happened from 1900 to early 1960s time-period.

In Europe: The only region world-wide that has stored data all the way from 1800s.
The infant mortality rate is quite stable during 1800-1875. 
Slightly decreased and then slightly increased. Yet, drastically decreased 
during 1875-1960s. Then slowly decreased after 1960s. 
```{r echo=TRUE, message=FALSE, warning=FALSE}
library(RSQLite)
library(dbplyr)
library(DBI)
library(ggplot2)

setwd("~/Desktop/5110/HM3/ddf--gapminder--systema_globalis-master")
library("readr")
country_entities <- read_csv("ddf--entities--geo--country.csv")


setwd("~/Desktop/5110/HM3/ddf--gapminder--systema_globalis-master/countries-etc-datapoints")
infant_mortality_rate <- read_csv(
  "ddf--datapoints--infant_mortality_rate_per_1000_births--by--geo--time.csv")

con <- dbConnect(SQLite(), ":memory:")
dbWriteTable(con, "infant_mortality_rate", infant_mortality_rate)
dbWriteTable(con, "country_entities", country_entities)

joined_infant_mortality_dt <-dbGetQuery(con, "SELECT * 
                                        FROM infant_mortality_rate 
                                        JOIN country_entities 
                                        ON geo=country 
                                        ORDER BY time DESC")

dbWriteTable(con, "joined_infant_mortality_dt", joined_infant_mortality_dt)

ggplot(data=joined_infant_mortality_dt, 
       mapping=aes(x=time, 
                   y=infant_mortality_rate_per_1000_births, 
                   fill=world_4region)) + 
  geom_point(position="jitter") +
  geom_smooth(method = 'gam') + 
  labs(x="Years", 
       y="Infant Mortality Rate Per 1000 Births", title=
"The Overall World's Infant Mortality Rate Decreases over time period (1800-2015)") +
  facet_wrap(~world_4region) +
  theme_minimal() 
  
```

## Problem 5

In Europe: The graph shows consistancy of the trend that when life expectancy 
increases, infant mortality rate decreases over time.

In America: Same as the trend in Europe.

In Asia: Both infant mortality rate(IMR) and life expectancy years(LEY) increase when 
life-expectancy is around 0-30. Then IMR decreased while LEY increased.

In Africa:Both infant mortality rate(IMR) and life expectancy years(LEY) increase when 
life-expectancy is around 0-35. Then IMR decreased while LEY increased.

```{r echo=TRUE, message=FALSE, warning=FALSE}
setwd("~/Desktop/5110/HM3/ddf--gapminder--systema_globalis-master/countries-etc-datapoints")
life_expectancy_years <- read_csv(
  "ddf--datapoints--life_expectancy_years--by--geo--time.csv")
dbWriteTable(con, "life_expectancy_years", life_expectancy_years)

library(dplyr)

lifeE_infantM <- dbGetQuery(con, "SELECT DISTINCT
joined_infant_mortality_dt.geo,
joined_infant_mortality_dt.time,
joined_infant_mortality_dt.world_4region,
joined_infant_mortality_dt.infant_mortality_rate_per_1000_births,
life_expectancy_years.life_expectancy_years
FROM joined_infant_mortality_dt
INNER JOIN life_expectancy_years
WHERE life_expectancy_years.geo=joined_infant_mortality_dt.geo
AND life_expectancy_years.time=joined_infant_mortality_dt.time
")

dbWriteTable(con, "lifeE_infantM", lifeE_infantM)


ggplot(lifeE_infantM, mapping=aes(x=life_expectancy_years, 
                                  y=infant_mortality_rate_per_1000_births,
                                  fill=world_4region)) +
  geom_point() + geom_smooth() + 
  labs(x="Life Expectancy Years", 
       y="Infant Mortality Rate Per 1000 Births", 
       title=
"The higher life-expectancy the lower infant mortality rate over time period (1800-2015)") + theme_minimal()
```

