---
title: "HM4"
author: "Mu Cheng"
date: "3/15/2021"
output: pdf_document
---

## Problem1

Step1: Call the Original Dataframe of the Life_Expectancy
```{r echo=TRUE, message=FALSE, warning=FALSE}
getwd()
setwd("~/Desktop/5110/HM4/ddf--gapminder--systema_globalis-master/countries-etc-datapoints")
library("readr")
library(ggplot2)
library(RSQLite)
library(dbplyr)
library(DBI)

life_expectancy_ori <- read_csv("ddf--datapoints--life_expectancy_years--by--geo--time.csv")
life_expectancy_ori <- na.omit(life_expectancy_ori)


setwd("~/Desktop/5110/HM4/ddf--gapminder--systema_globalis-master/countries-etc-datapoints")
infant_mortality_ori <- read_csv(
  "ddf--datapoints--infant_mortality_rate_per_1000_births--by--geo--time.csv")
infant_mortality_ori <- na.omit(infant_mortality_ori)


setwd("~/Desktop/5110/HM4/ddf--gapminder--systema_globalis-master/countries-etc-datapoints")
murder_ori <- read_csv(
  "ddf--datapoints--murder_per_100000_people--by--geo--time.csv")
murder_ori <- na.omit(murder_ori)


setwd("~/Desktop/5110/HM4/ddf--gapminder--systema_globalis-master/countries-etc-datapoints")
GDP_inflation_ori <- read_csv(
  "ddf--datapoints--total_gdp_us_inflation_adjusted--by--geo--time.csv")
GDP_inflation_ori <- na.omit(GDP_inflation_ori)


setwd("~/Desktop/5110/HM4/ddf--gapminder--systema_globalis-master/countries-etc-datapoints")
MD_ori <- read_csv(
  "ddf--datapoints--medical_doctors_per_1000_people--by--geo--time.csv")
MD_ori <- na.omit(MD_ori)


setwd("~/Desktop/5110/HM4/ddf--gapminder--systema_globalis-master/countries-etc-datapoints")
poverty_rate_ori <- read_csv(
  "ddf--datapoints--poverty_percent_people_below_550_a_day--by--geo--time.csv")
poverty_rate_ori <- na.omit(poverty_rate_ori)


con <- dbConnect(SQLite(), ":memory:")
dbWriteTable(con, "life_expectancy_ori", life_expectancy_ori)
dbWriteTable(con, "infant_mortality_ori", infant_mortality_ori)
dbWriteTable(con, "murder_ori", murder_ori)
dbWriteTable(con, "GDP_inflation_ori", GDP_inflation_ori)
dbWriteTable(con, "MD_ori", MD_ori)
dbWriteTable(con, "poverty_rate_ori", poverty_rate_ori)

LE_data1 <-dbGetQuery(con, "SELECT DISTINCT
infant_mortality_ori.geo,
infant_mortality_ori.time,
infant_mortality_ori.infant_mortality_rate_per_1000_births,
life_expectancy_ori.life_expectancy_years
FROM life_expectancy_ori
INNER JOIN infant_mortality_ori
WHERE infant_mortality_ori.geo = life_expectancy_ori.geo
AND infant_mortality_ori.time=life_expectancy_ori.time")
dbWriteTable(con, "LE_data1", LE_data1)


LE_data2 <-dbGetQuery(con, "SELECT DISTINCT
murder_ori.geo,
murder_ori.time,
murder_ori.murder_per_100000_people,
GDP_inflation_ori.total_gdp_us_inflation_adjusted
FROM GDP_inflation_ori
INNER JOIN murder_ori
WHERE murder_ori.geo=GDP_inflation_ori.geo
AND murder_ori.time=GDP_inflation_ori.time")
dbWriteTable(con, "LE_data2", LE_data2)


LE_data3 <-dbGetQuery(con, "SELECT DISTINCT
MD_ori.geo,
MD_ori.time,
MD_ori.medical_doctors_per_1000_people,
poverty_rate_ori.poverty_percent_people_below_550_a_day
FROM poverty_rate_ori
INNER JOIN MD_ori
WHERE MD_ori.geo = poverty_rate_ori.geo
AND MD_ori.time=poverty_rate_ori.time")
dbWriteTable(con, "LE_data3", LE_data3)


LE_data4 <- dbGetQuery(con, "SELECT
LE_data1.geo,
LE_data1.time,
LE_data1.infant_mortality_rate_per_1000_births,
LE_data1.life_expectancy_years,
LE_data2.murder_per_100000_people,
LE_data2.total_gdp_us_inflation_adjusted
FROM LE_data1
JOIN LE_data2
WHERE LE_data2.geo=LE_data1.geo
AND LE_data2.time=LE_data1.time
")
dbWriteTable(con, "LE_data4", LE_data4)

LE_data <- dbGetQuery(con, "SELECT
LE_data4.geo,
LE_data4.time,
LE_data4.infant_mortality_rate_per_1000_births,
LE_data4.life_expectancy_years,
LE_data4.murder_per_100000_people,
LE_data4.total_gdp_us_inflation_adjusted,
LE_data3.medical_doctors_per_1000_people,
LE_data3.poverty_percent_people_below_550_a_day
FROM LE_data4
JOIN LE_data3
WHERE LE_data3.geo=LE_data4.geo
AND LE_data3.time=LE_data4.time
")
dbWriteTable(con, "LE_data", LE_data)
```
le1: life_expectancy_years against infant_mortality_rate_per_1000_births.
Log10 transformation on both variables.
```{r echo=TRUE, message=FALSE, warning=FALSE}
le1<- ggplot(LE_data, mapping=aes(x=log10(life_expectancy_years), 
                                  y=log10(infant_mortality_rate_per_1000_births))) +
  geom_point(alpha=0.1) + geom_smooth(method = 'loess', formula='y ~ x') + geom_smooth(method="lm", color="red") +
  labs(x="Life Expectancy Years", 
       y="Infant Mortality Rate Per 1000 Births", 
       title=
"The higher life-expectancy the lower infant mortality rate over time period (1800-2015)") + 
  theme_minimal()
le1
```

le2: life_expectancy_years against murder_per_100000_people.
Log10 transformations on both variables.
```{r echo=TRUE, message=FALSE, warning=FALSE}
le2<- ggplot(LE_data, mapping=aes(x=log10(life_expectancy_years), 
                                  y=log10(murder_per_100000_people))) +
  geom_point(alpha=0.1) + geom_smooth(method = 'loess', formula='y ~ x') +  geom_smooth(method="lm", color="red") +
  labs(x="Life Expectancy Years", 
       y="murder_per_100000_people", 
       title=
"As life expectancy increases in years, murder rate goes down during (1800-2015)") + theme_minimal()
le2
```

le3: life_expectancy_years against total_gdp_us_inflation_adjusted.
Log10 transformations on x variable.
```{r echo=TRUE, message=FALSE, warning=FALSE}
le3 <- ggplot(LE_data, mapping=aes(x=log10(life_expectancy_years), 
                                  y=total_gdp_us_inflation_adjusted)) +
  geom_point(alpha=0.1) + geom_smooth(method = 'loess', formula='y ~ x') +  geom_smooth(method="lm", color="red") +
  labs(x="Life Expectancy Years", 
       y="total_gdp_us_inflation_adjusted", 
       title=
"No Specific relation is found between life expectancy and gdp inflation between (1800-2015)") + theme_minimal()
le3
```

le4: life_expectancy_years against medical_doctors_per_1000_people.
Log10 transformations on both variables.
```{r echo=TRUE, message=FALSE, warning=FALSE}
le4 <- ggplot(LE_data, mapping=aes(x=log10(life_expectancy_years), 
                                  y=log10(medical_doctors_per_1000_people))) +
  geom_point(alpha=0.1) + geom_smooth(method = 'loess', formula='y ~ x') +  geom_smooth(method="lm", color="red") +
  labs(x="Life Expectancy Years", 
       y="medical_doctors_per_1000_people", 
       title=
"More medical doctors continue to be needed as life expectancy increases during (1800-2015)") + theme_minimal()
le4
```

le5: life_expectancy_years against poverty_percent_people_below_550_a_day.
Log10 transformations on x variable.
```{r echo=TRUE, message=FALSE, warning=FALSE}

le5 <- ggplot(LE_data, mapping=aes(x=log10(life_expectancy_years), 
                                  y=poverty_percent_people_below_550_a_day)) +
  geom_point(alpha=0.1) + geom_smooth(method = 'loess', formula='y ~ x') + 
  geom_smooth(method="lm", color="red") +
  labs(x="Life Expectancy Years", 
       y="poverty_percent_people_below_550_a_day", 
       title=
"Poverty rate in general decreases as life expectancy increases during (1800-2015)") + 
  theme_minimal()
le5
```

## Problem2

Based on previous graphs, the response variable, Murder_per_10000_people, has the 
best linearity among all the plots. Therefore, Murder_per_10000_people is 
utilized here as a predictor.

Residual graph demonstrates a clear pattern(increasing line) when plotting
without log transformation, with r squared value around 0.3115. It's in violation
of model assumptions(linearity) Therefore, adjustment of adding log10() to both 
variables is made. Then the value of r squared is better with 0.6112. 
After making the adjustments, the plotted residual graph with a better 
model (fit1) is then normal in demonstration.

Step 1. Compare better fit vs. bad fit's summary stats.
```{r echo=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(modelr)
library(broom)

murder_graph <- ggplot(LE_data, aes(x=log10(life_expectancy_years), 
                                    y=log10(murder_per_100000_people))) +
  geom_point(alpha=0.1) + geom_smooth() + geom_smooth(method="lm", color="red") +
  labs(x="Life Expectancy in Years", 
       y="Number of murder per 10000 people",
    title="As life expectancy increases in years, murder rate goes down during (1800-2015)") +
  geom_smooth(method="lm", color="red") + theme_minimal()
murder_graph
  
fit1 <- lm(log10(life_expectancy_years) ~ log10(murder_per_100000_people), 
           data=LE_data)
summary(fit1)
glance(fit1)

fitbad <- lm(life_expectancy_years ~ murder_per_100000_people, 
           data=LE_data)
summary(fitbad)
glance(fitbad)

```

Step 2.
Compare both fit and bad models on both residual plots, then check the
normality of the two residuals by using qq_plot, finally check the residual
outlier with olsrr package.

```{r echo=TRUE, message=FALSE, warning=FALSE}
library("olsrr")
bad_residual <- LE_data %>%
  add_residuals(fitbad, "resid") %>%
  ggplot(aes(x=log10(life_expectancy_years))) +
  geom_point(aes(y=resid), alpha=0.1) +
  labs(x="life expectancy in years", y="residuals") +
  theme_minimal()
bad_residual

bad_residual_qq <- LE_data %>%
  add_residuals(fitbad, "resid") %>%
  ggplot(aes(sample=resid)) +
  geom_qq() +
  theme_minimal()
bad_residual_qq

bad_residual_box <- ols_plot_resid_box(fitbad)
bad_residual_box

better_residual <- LE_data %>%
  add_residuals(fit1, "resid") %>%
  ggplot(aes(x=log10(life_expectancy_years))) +
  geom_point(aes(y=resid), alpha=0.1) +
  labs(x="life expectancy in years", y="residuals") +
  theme_minimal()
better_residual

better_residual_qq <- LE_data %>%
  add_residuals(fit1, "resid") %>%
  ggplot(aes(sample=resid)) +
  geom_qq() + labs(y="sample", x="theoretical")
  theme_minimal()
better_residual_qq

better_residual_box <- ols_plot_resid_box(fit1)
better_residual_box
```

Removing outliers. TO BE REMOVED. Not so much of the difference.
```{r echo=TRUE, message=FALSE, warning=FALSE}
better_residual_removedOutlier <- fit1[!fit1 %in% better_residual_box$out]
boxplot(x=better_residual_removedOutlier$residuals,
        y=better_residual_removedOutlier$residuals)

```   

## Problem3

Using total_gdp_us_inflation_adjusted as the one additional predictor after 
comparing the 4 different predictors. The adjusted R^2 increases to 0.6311 
from 0.6112.
```{r echo=TRUE, message=FALSE, warning=FALSE}
pred_infant <- LE_data %>%
  add_residuals(fit1, "resid") %>%
  ggplot(aes(x=infant_mortality_rate_per_1000_births, y=resid)) +
  geom_point() + geom_smooth() +
  labs(x="infant_mortality_rate_per_1000_births", y="Residuals") +
  theme_minimal()
pred_infant

pred_gdp <- LE_data %>%
  add_residuals(fit1, "resid") %>%
  ggplot(aes(x=total_gdp_us_inflation_adjusted, y=resid)) +
  geom_point() + geom_smooth() +
  labs(x="total_gdp_us_inflation_adjusted", y="Residuals") +
  theme_minimal()
pred_gdp

pred_dc <- LE_data %>%
  add_residuals(fit1, "resid") %>%
  ggplot(aes(x=medical_doctors_per_1000_people, y=resid)) +
  geom_point() + geom_smooth() +
  labs(x="medical_doctors_per_1000_people", y="Residuals") +
  theme_minimal()
pred_dc

pred_poverty <- LE_data %>%
  add_residuals(fit1, "resid") %>%
  ggplot(aes(x=poverty_percent_people_below_550_a_day, y=resid)) +
   geom_point() + geom_smooth() +
  labs(x="poverty_percent_people_below_550_a_day", y="Residuals") +
  theme_minimal()
pred_poverty

```



```{r echo=TRUE, message=FALSE, warning=FALSE}
fit2 <- lm(log10(life_expectancy_years) ~ log10(murder_per_100000_people) +
             total_gdp_us_inflation_adjusted, data=LE_data)
summary(fit2)

better_residual <- LE_data %>%
  add_residuals(fit2, "resid") %>%
  ggplot(aes(x=log10(life_expectancy_years))) +
  geom_point(aes(y=resid), alpha=0.1) +
  labs(x="life expectancy in years", y="residuals") +
  theme_minimal()
better_residual

better_residual_qq <- LE_data %>%
  add_residuals(fit2, "resid") %>%
  ggplot(aes(sample=resid)) +
  geom_qq() + labs(y="sample", x="theoretical", title="Normal Residual Graph")
  theme_minimal()
better_residual_qq
```

## Problem4

The better model from problem3 has a smaller cv RMSE at 0.01565241 whereas
the original model has a larger cv RMSE at 0.0161054. It's not surprising as
we know the more variables we add the better model will show a smaller CV RMSE.

```{r echo=TRUE, message=FALSE, warning=FALSE}
LE_data_cv <- crossv_kfold(LE_data, 10)
LE_data_cv

cv_better_model <- LE_data_cv %>%
  mutate(fit = purrr::map(train, 
                          ~lm(log10(life_expectancy_years) ~ 
                                log10(murder_per_100000_people) +
             total_gdp_us_inflation_adjusted, data=.)),
         rmse=purrr::map2_dbl(fit, test, ~rmse(.x, .y)))
cv_better_model

cv_original <- LE_data_cv %>%
  mutate(fit = purrr::map(train, 
                          ~lm(log10(life_expectancy_years) ~ 
                                log10(murder_per_100000_people), data=.)),
         rmse=purrr::map2_dbl(fit, test, ~rmse(.x, .y)))
cv_original

mean(cv_better_model$rmse)
mean(cv_original$rmse)
```

## Problem5

Partitioning
```{r echo=TRUE, message=FALSE, warning=FALSE}
set.seed(10)
LE_data_part <- resample_partition(LE_data,
                                   p=c(train=0.5,
                                       valid=0.25,
                                       test=0.25))
LE_data_part

step1 <- function(response, predictors, candidates, partition)
{
  rhs <- paste0(paste0(predictors, collapse="+"), "+", candidates)
  formulas <- lapply(paste0(response, "~", rhs), as.formula)
  rmses <- sapply(formulas,
                  function(fm) rmse(lm(fm, data=partition$train),
                                    data=partition$valid))
  names(rmses) <- candidates
  attr(rmses, "best") <- rmses[which.min(rmses)]
  rmses
}

model <- NULL
```


step1
```{r echo=TRUE, message=FALSE, warning=FALSE}
preds <- "1"
cands <- c("log10(infant_mortality_rate_per_1000_births)",
           "log10(murder_per_100000_people)", 
           "total_gdp_us_inflation_adjusted",
           "log10(medical_doctors_per_1000_people)",
           "poverty_percent_people_below_550_a_day")
s1 <- step1("log10(life_expectancy_years)", preds, cands, LE_data_part)

model <- c(model, attr(s1, "best"))
s1
```

Step2 adding log10(infant_mortality_rate_per_1000_births)
```{r echo=TRUE, message=FALSE, warning=FALSE}
preds <- "log10(infant_mortality_rate_per_1000_births)"
cands <- c("log10(murder_per_100000_people)", 
           "total_gdp_us_inflation_adjusted",
           "log10(medical_doctors_per_1000_people)",
           "poverty_percent_people_below_550_a_day")
s2 <- step1("log10(life_expectancy_years)", preds, cands, LE_data_part)

model <- c(model, attr(s2, "best"))
s2
```
Step 3 Adding "log10(murder_per_100000_people)"
```{r echo=TRUE, message=FALSE, warning=FALSE}
preds <- c("log10(infant_mortality_rate_per_1000_births)", 
"log10(murder_per_100000_people)")
cands <- c("total_gdp_us_inflation_adjusted",
           "log10(medical_doctors_per_1000_people)",
           "poverty_percent_people_below_550_a_day")
s3 <- step1("log10(life_expectancy_years)", preds, cands, LE_data_part)

model <- c(model, attr(s3, "best"))
s3
```
Step 4 Adding total_gdp_us_inflation_adjusted 
```{r echo=TRUE, message=FALSE, warning=FALSE}
preds <- c("log10(infant_mortality_rate_per_1000_births)", 
"log10(murder_per_100000_people)", "total_gdp_us_inflation_adjusted")
cands <- c("log10(medical_doctors_per_1000_people)",
           "poverty_percent_people_below_550_a_day")
s4 <- step1("log10(life_expectancy_years)", preds, cands, LE_data_part)

model <- c(model, attr(s4, "best"))
s4
```

Step 5 Adding poverty_percent_people_below_550_a_day.
Then stop at step 5 due to the increasing RMSE value of 
log(medical_doctors_per_1000_people).
```{r echo=TRUE, message=FALSE, warning=FALSE}
preds <- c("log10(infant_mortality_rate_per_1000_births)", 
"log10(murder_per_100000_people)", "total_gdp_us_inflation_adjusted",
"poverty_percent_people_below_550_a_day")
cands <- c("log10(medical_doctors_per_1000_people)")
s5 <- step1("log10(life_expectancy_years)", preds, cands, LE_data_part)

model <- c(model, attr(s5, "best"))
s5
```

Plotting
```{r echo=TRUE, message=FALSE, warning=FALSE}
step_model <- tibble(index=seq_along(model),
                     variable=factor(names(model), levels=names(model)),
                     RMSE=model)

ggplot(step_model, aes(y=RMSE)) +
  geom_point(aes(x=variable)) +
  geom_line(aes(x=index)) +
  labs(title="Stepwise model selection") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle=90, vjust=0.5,  hjust=1))
```