---
title: "R Notebook"
output: html_notebook
---


#Embbed ERD Diagram for Part1
```{r out.width="300px"}
url <- "https://imgur.com/a/GbsqAqc"
knitr::include_graphics(url)
```

<div style="width: 640px; height: 480px; margin: 10px; position: relative;"><iframe allowfullscreen frameborder="0" style="width:640px; height:480px" src="https://lucid.app/documents/embeddedchart/f3b60aca-ca3e-46bd-b8f7-88f4a31c5ccc" id="BH7A4gCUen6t"></iframe></div>


#Creating Journal database
```{r}
library(RSQLite)

fpath = "~/Desktop/NEU/5200DBMS/Database/"
dbfile = "Journal.sqlitedb"
jcon <- dbConnect(RSQLite::SQLite(), paste0(fpath, dbfile))
```


#Creating Tables in Journal.sqlitedb Database
# 3 Tables are stored in this database
# Articles
# Journals
# Authors

# Special note in question 2, the publication date is equivalent to PubMedPubDate PubStatus="pubmed"


```{sql connection=jcon}
DROP TABLE IF EXISTS Articles
```

```{sql connection=jcon}
CREATE TABLE Articles (
	article_id NUMERIC NOT NULL,
	journal_id NUMERIC NOT NULL,
  article_title TEXT NOT NULL,
  date_created DATE NOT NULL,
  PRIMARY KEY (article_id),
  CONSTRAINT journal_id_fk FOREIGN KEY (journal_id) REFERENCES Journals (journal_id)
)
```

```{sql connection=jcon}
DROP TABLE IF EXISTS Journals
```

#journal_pubDate(PubMedPubDate PubStatus="pubmed")
```{sql connection=jcon}
CREATE TABLE Journals (
	journal_id NUMERIC NOT NULL,
  journal_title TEXT NOT NULL,
  volume NUMERIC NOT NULL,
  issue NUMERIC NOT NULL,
  pubDate DATE NOT NULL,
  PRIMARY KEY (journal_id)
)
```

```{sql connection=jcon}
DROP TABLE IF EXISTS Authors
```

```{sql connection=jcon}
CREATE TABLE Authors (
	author_id NUMERIC NOT NULL,
  foreName TEXT NOT NULL,
  lastName TEXT NOT NULL,
  initials TEXT NOT NULL,
  affiliation TEXT NOT NULL,
  PRIMARY KEY (author_id)
)
```


```{r}
dbListTables(jcon)
```

#Part 1: Loading Tables with data information from XML

```{r}
library(XML)

library(RSQLite)
library(DBI)
library(knitr)


path <- "~/Desktop/"
fn <- "pubmed_sample.xml"
fpn<- paste0(path, fn)

xmlDOM <- xmlParse(file = fpn)
```

```{r}
r <- xmlRoot(xmlDOM)
numArticle<- xmlSize(r)
numArticle

```

#Preparing values to enter the Author table
```{r}
fName_xpath <- '//PubmedArticle/MedlineCitation/Article/AuthorList/Author/ForeName'
fNames <- xpathSApply(xmlDOM,fName_xpath, xmlValue)
#fNames

lName_xpath <- '//PubmedArticle/MedlineCitation/Article/AuthorList/Author/LastName'
lNames <- xpathSApply(xmlDOM,lName_xpath, xmlValue)

initial_xpath <- '//PubmedArticle/MedlineCitation/Article/AuthorList/Author/Initials'
initial <- xpathSApply(xmlDOM,initial_xpath, xmlValue)

affiliation_xpath <- '//PubmedArticle/MedlineCitation/Article/AuthorList/Author/Affiliation'
affiliation <- xpathSApply(xmlDOM,affiliation_xpath, xmlValue)

```

#Creating a subset dataframe for authors who have affiliations called Author_with_affi.df
#Loading data to Author_with_affi.df 
```{r}
Author_with_affi.df <- data.frame(Author_ForeName=character(),
                                  Author_LastName=character(),
                                  Author_Initial=character(),
                                  Author_Affiliation=character(),
                            stringsAsFactors = F)

for (m in 1:length(affiliation)){

fn <- fNames[m]
ln <- lNames[m]
ini <- initial[m]
affi <- affiliation[m]

Author_with_affi.df[m,1] <- fn
Author_with_affi.df[m,2] <- ln
Author_with_affi.df[m,3] <- ini
Author_with_affi.df[m,4] <- affi
}

Author_with_affi.df

```


#Loading data to All_Author.df for all authors in the xml file
#Generating author ids (no duplicated authors)
```{r}
All_Author.df <- data.frame(Author_ID=integer(),
                            authorFN=character(),
                      authorLN=character(),
                      authorINI=character(),
                      stringsAsFactors = F)

for (m in 1:length(lName)){
first_name <- fNames[m]
last_name <- lNames[m]
inital <- initial[m]

All_Author.df[m,2] <- first_name
All_Author.df[m,3] <- last_name
All_Author.df[m,4] <- inital
}
All_Author.df <- All_Author.df[!duplicated(All_Author.df),]
All_Author.df$Author_ID <- seq(from = 1, to = nrow(All_Author.df)) + 100
All_Author.df
```
#Complete Author.df by joining All_Author.df and Author_with_affi.df
```{r}
Author.df <- merge(x=Author_with_affi.df, y=All_Author.df,
                        by.x=c("Author_ForeName", "Author_LastName", "Author_Initial"),by.y=c("authorFN", "authorLN", "authorINI"),
                        all.y=TRUE)
Author.df 

numAffiAuwAll <- nrow(Author.df)
numAffiAuwAll

numAffiAu <- sum(is.na(Author.df$Author_Affiliation))
numAffiAu 

Author.df <- subset(Author.df, select = c(5, 1, 2, 3, 4))
Author.df
```


#Preparing values to enter the Journal table
```{r}
jID_xpath <- '//PubmedArticle/MedlineCitation/MedlineJournalInfo/NlmUniqueID'
jID <- xpathSApply(xmlDOM,jID_xpath, xmlValue)
#jID

jTitle_xpath <- '//PubmedArticle/MedlineCitation/Article/Journal/Title'
jTitle <- xpathSApply(xmlDOM,jTitle_xpath, xmlValue)
#jTitle

jVolume_xpath <- '//PubmedArticle/MedlineCitation/Article/Journal/JournalIssue/Volume'
jVolume <- xpathSApply(xmlDOM,jVolume_xpath, xmlValue)
#jVolume

jIssue_xpath <- '//PubmedArticle/MedlineCitation/Article/Journal/JournalIssue/Issue'
jIssue <- xpathSApply(xmlDOM,jIssue_xpath, xmlValue)
#jIssue

jDateYear_xpath <- "//PubmedArticle/PubmedData/History/PubMedPubDate[@PubStatus='medline']/Year"
jDateYear<- xpathSApply(xmlDOM,jDateYear_xpath, xmlValue)
#jDateYear

jDateMonth_xpath <- "//PubmedArticle/PubmedData/History/PubMedPubDate[@PubStatus='medline']/Month"
jDateMonth<- xpathSApply(xmlDOM,jDateMonth_xpath, xmlValue)
#jDateMonth

jDateDay_xpath <- "//PubmedArticle/PubmedData/History/PubMedPubDate[@PubStatus='medline']/Day"
jDateDay<- xpathSApply(xmlDOM,jDateDay_xpath, xmlValue)
#jDateDay
```



#Loading data for journal.df
```{r}
journal.df <- data.frame(Journal_ID=integer(),
                         Journal_Title=character(),
                         Journal_Volume=integer(),
                         Journal_Issue=integer(),
                         Journal_PubYear=integer(),
                         Journal_PubMonth=integer(),
                         Journal_PubDay=integer(),
                         stringsAsFactors = F)

for (m in 1:length(jID)){
journal.df[m,1] <- jID[m]
journal.df[m,2] <- jTitle[m]
journal.df[m,3] <- jVolume[m]
journal.df[m,4] <- jIssue[m]
journal.df[m,5] <- jDateYear[m]
journal.df[m,6] <- jDateMonth[m]
journal.df[m,7] <- jDateDay[m]
}

#journal.df$Journal_PubDate <- as.Date(journal.df$Journal_PubDate)
journal.df$Date <- paste(journal.df$Journal_PubYear, 
                         journal.df$Journal_PubMonth, 
                         journal.df$Journal_PubDay, 
                         sep="-")
#Convert to the correct data types
journal.df$Journal_ID <- as.numeric(journal.df$Journal_ID)
journal.df$Journal_Volume <- as.numeric(journal.df$Journal_Volume)
journal.df$Journal_Issue <- as.numeric(journal.df$Journal_Issue)
journal.df$Date <- as.Date(journal.df$Date, format="%Y-%m-%d")

journal.df <- subset(journal.df, select = c(1,2,3,4,8))
journal.df <- journal.df[!duplicated(journal.df),]
journal.df
```
#Problem occurs due to automatic removal of leading zeros of the journal ID
#Adding zeros to the journal_ID
#2,3,6,9,10,16
```{r}
journal.df$Journal_ID[2] <- paste(0, journal.df$Journal_ID[2], sep = '')
journal.df$Journal_ID[3] <- paste(0, journal.df$Journal_ID[3], sep = '')
journal.df$Journal_ID[6] <- paste(0, journal.df$Journal_ID[6], sep = '')
journal.df$Journal_ID[9] <- paste(0, journal.df$Journal_ID[9], sep = '')
journal.df$Journal_ID[10] <- paste(0, journal.df$Journal_ID[10], sep = '')
journal.df$Journal_ID[16] <- paste(0, journal.df$Journal_ID[16], sep = '')
journal.df$Journal_ID[6] <- paste(0, journal.df$Journal_ID[6], sep = '')
journal.df
```


#Preparing values to enter the Article table
```{r}
articleID_xpath <- "//PubmedArticle/PubmedData/ArticleIdList/ArticleId[@ IdType='pubmed']"
articleID <- xpathSApply(xmlDOM,articleID_xpath, xmlValue)
#articleID

title_xpath <- '//PubmedArticle/MedlineCitation/Article/ArticleTitle'
titles <- xpathSApply(xmlDOM,title_xpath, xmlValue)
#titles

dateCreatedYear_xpath <- "//PubmedArticle/MedlineCitation/DateCreated/Year"
dateCreatedYear <- xpathSApply(xmlDOM,dateCreatedYear_xpath, xmlValue)
#dateCreatedYear

dateCreatedMonth_xpath <- "//PubmedArticle/MedlineCitation/DateCreated/Month"
dateCreatedMonth <- xpathSApply(xmlDOM,dateCreatedMonth_xpath, xmlValue)
#dateCreatedMonth

dateCreatedDay_xpath <- "//PubmedArticle/MedlineCitation/DateCreated/Day"
dateCreatedDay <- xpathSApply(xmlDOM,dateCreatedDay_xpath, xmlValue)
#dateCreatedDay
```

#Loading data for article.df
```{r}
article_with_DaysInfo.df <- data.frame(Article_ID=integer(),
                         Journal_ID=integer(),
                         Article_Title=character(),
                         Date_Year=integer(),
                         Date_Month=integer(),
                         Date_Day=integer(),
                         stringsAsFactors = F)

for (m in 1:length(jID)){
article_with_DaysInfo.df[m,1] <- articleID[m]
article_with_DaysInfo.df[m,2] <- jID[m]
article_with_DaysInfo.df[m,3] <- titles[m]
article_with_DaysInfo.df[m,4] <- dateCreatedYear[m]
article_with_DaysInfo.df[m,5] <- dateCreatedMonth[m]
article_with_DaysInfo.df[m,6] <- dateCreatedDay[m]
}


article_with_DaysInfo.df$Date_Created <- paste(article_with_DaysInfo.df$Date_Year, 
                         article_with_DaysInfo.df$Date_Month, 
                         article_with_DaysInfo.df$Date_Day, 
                         sep="-")
#Convert to the correct data types
article_with_DaysInfo.df$Article_ID <- as.integer(article_with_DaysInfo.df$Article_ID)
article_with_DaysInfo.df$Journal_ID <- as.integer(article_with_DaysInfo.df$Journal_ID)
article_with_DaysInfo.df$Date_Created <- as.Date(article_with_DaysInfo.df$Date_Created, format="%Y-%m-%d")
article_with_DaysInfo.df
#article.df_with_DaysInfo.df for creating a Created_Day.df dimensional table later in part 2.
#Now just focus on article.df
article.df <- subset(article.df_with_DaysInfo.df, select = c(1,2,3,7))
article.df
```
#Problem occurs due to automatic removal of leading zeros of the journal ID
#Adding zeros to the journal_ID
#2,3,6,9,10,16
```{r}
#article.df$Journal_ID[2] <- paste(0, article.df$Journal_ID[2], sep = '')
#article.df$Journal_ID[3] <- paste(0, article.df$Journal_ID[3], sep = '')
#article.df$Journal_ID[6] <- paste(0, article.df$Journal_ID[6], sep = '')
#article.df$Journal_ID[9] <- paste(0, article.df$Journal_ID[9], sep = '')
#article.df$Journal_ID[10] <- paste(0, article.df$Journal_ID[10], sep = '')
#article.df$Journal_ID[16] <- paste(0, article.df$Journal_ID[16], sep = '')
#article.df$Journal_ID[6] <- paste(0, article.df$Journal_ID[6], sep = '')
#article.df
```

```{r}
#article.df
#journal.df
#Author.df
```
#Populating tables in journal database
```{r}
dbWriteTable(jcon, "Articles", article.df, overwrite = T, row.names =F)
```
#Journals Table

```{r}
journal.df$Date <- as.character(journal.df$Date)
#journal.df
```

```{r}
dbWriteTable(jcon, "Journals", journal.df, overwrite = T, row.names =F)
```
#Authors Table
```{r}
dbWriteTable(jcon, "Authors", Author.df, overwrite = T, row.names =F)
```
#Show first 10 rows of the Articles Table
```{sql connection=jcon}
select * from Articles limit 10
```
#Show first 10 rows of the Authors Table
```{sql connection=jcon}
select * from Authors limit 10
```


#Show first 10 rows of the Journals Table
```{sql connection=jcon}
select * from Journals limit 10
```


#PART2. Create Star/Snowflake Schema


#Embbed ERD Diagram for Part2
```{r out.width="300px"}
url <- "https://imgur.com/a/ofgnwuK"
knitr::include_graphics(url)
```

<div style="width: 640px; height: 480px; margin: 10px; position: relative;"><iframe allowfullscreen frameborder="0" style="width:640px; height:480px" src="https://lucid.app/documents/embeddedchart/624b9f66-8f50-4472-b4a7-c26d913d6f69" id="cK7AYlGc0iQc"></iframe></div>



#Step1: Creating Tables

#Creating database tables for Journal_fact, Number_Article.dim, 
#Publication_Date.dim, Created_Date.dim, and Days_Elapsed.dim

#Fact Table
```{sql connection=jcon}
DROP TABLE IF EXISTS Journal_fact
```

```{sql connection=jcon}
CREATE TABLE Journal_fact (
	journal_id NUMERIC NOT NULL,
	Pub_ID TEXT NOT NULL,
	Created_ID TEXT NOT NULL,
	ED_Key TEXT NOT NULL,
	Date_Created TEXT NOT NULL,
	Date_Published TEXT NOT NULL,
	Quarter_Published TEXT NOT NULL,
	Days_Elapsed NUMERIC NOT NULL,
	AVG_ED_Q NUMERIC NOT NULL,
	AVG_ED_Y NUMERIC NOT NULL,
	Number_of_Articles NUMERIC NOT NULL,
  PRIMARY KEY (journal_id,Pub_ID, Created_ID, ED_Key),
  CONSTRAINT journal_id_fk FOREIGN KEY (journal_id) REFERENCES Number_Article_dim (Journal_ID),
  CONSTRAINT pub_id_fk FOREIGN KEY (Pub_ID) REFERENCES Publication_Date_dim (Pub_ID),
  CONSTRAINT created_id_fk FOREIGN KEY (Created_ID) REFERENCES Created_Date_dim (Created_ID),
  CONSTRAINT ed_key_fk FOREIGN KEY (ED_Key) REFERENCES Days_Elapsed_dim (ED_Key)
)
```

#Create Dimensional Table#1 Number_Article.dim
```{sql connection=jcon}
DROP TABLE IF EXISTS Number_Article_dim
```

```{sql connection=jcon}
CREATE TABLE Number_Article_dim (
	journal_id NUMERIC NOT NULL,
	number_of_articles NUMERIC NOT NULL,
  PRIMARY KEY (journal_id),
  CONSTRAINT journal_id_fk FOREIGN KEY (journal_id) REFERENCES Journal_fact (journal_id)
)
```

#Create Dimensional Table #2 Publication_Date.dim
```{sql connection=jcon}
DROP TABLE IF EXISTS Publication_Date_dim
```

```{sql connection=jcon}
CREATE TABLE Publication_Date_dim(
	Pub_ID TEXT NOT NULL,
	Date_Published TEXT NOT NULL,
	Year_Published TEXT NOT NULL,
	Month_Published TEXT NOT NULL,
	Day_Published TEXT NOT NULL,
	Date_Published_in_Format TEXT NOT NULL,
  PRIMARY KEY (Pub_ID),
  CONSTRAINT pub_id_fk FOREIGN KEY (Pub_ID) REFERENCES Journal_fact (Pub_ID)
)
```

#Create Dimensional Table#3 Created_Date.dim
```{sql connection=jcon}
DROP TABLE IF EXISTS Created_Date_dim
```

```{sql connection=jcon}
CREATE TABLE Created_Date_dim(
	Created_ID TEXT NOT NULL,
	Date_Created_in_Format TEXT NOT NULL,
	Year_Created TEXT NOT NULL,
	Month_Created TEXT NOT NULL,
	Day_Created TEXT NOT NULL,
  PRIMARY KEY (Created_ID),
  CONSTRAINT created_id_fk FOREIGN KEY (Created_ID) REFERENCES Journal_fact (Created_ID)
)
```

#Create Dimensional Table#4 Days_Elapsed.dim
```{sql connection=jcon}
DROP TABLE IF EXISTS Days_Elapsed_dim
```

```{sql connection=jcon}
CREATE TABLE Days_Elapsed_dim(
	ED_Key TEXT NOT NULL,
	Journal_ID NUMERIC NOT NULL,
	Date_Created TEXT NOT NULL,
	Date_Published TEXT NOT NULL,
	Day_Elapsed TEXT NOT NULL,
	Month_Published TEXT NOT NULL,
	Quarter_Published TEXT NOT NULL,
	Year_Published TEXT NOT NULL,
  PRIMARY KEY (ED_Key, Journal_ID),
  CONSTRAINT journal_id_fk FOREIGN KEY (Journal_ID) REFERENCES Journal_fact (journal_id),
  CONSTRAINT ED_Key_fk FOREIGN KEY (ED_Key) REFERENCES Journal_fact (ED_Key)
)
```


```{r}
dbListTables(jcon)
```


#Step2: Populating dataframes

#Create a dataframe named numArti.df
```{r}
#Creating numArti_with_duplicated_jID.df first.
journalID <- jID
numArti_with_duplicated_jID.df <- data.frame(Journal_ID=integer(),
                         Number_of_Articles=integer(),
                         stringsAsFactors = F)
for (i in 1:19) {
anArticle <- r[[i]]
article_value <- length(xmlValue(anArticle[[1]][[5]][[2]]))

numArti_with_duplicated_jID.df[i,1] <- journalID[i]
numArti_with_duplicated_jID.df[i,2] <- article_value
}

#Dealing with duplicated journal ID and named the new df as numArti.df
numArti.df <- aggregate(numArti_with_duplicated_jID.df$Number_of_Articles, by=list(numArti_with_duplicated_jID.df$Journal_ID), FUN=sum)
colnames(numArti.df) <- c("Journal_ID", "Number_of_Articles")
numArti.df 
```

#Create a dataframe named pubDate.df(for dimensional table) to store
#Pupblication Date information only

#Extracting pubDate data
```{r}
library(stringr)
pubDate_xpath <- "//PubmedArticle/MedlineCitation/Article/Journal/JournalIssue/PubDate"
pubDate <- xpathSApply(xmlDOM,pubDate_xpath, xmlValue)
pubDate <- str_replace_all(pubDate, fixed(" "), "")
pubDate
```

#substring pubDate into YYYY-MM-DD format
#  types of formats that are recognized
#Format1.Ending with number, eg. 2012Feb1
#Format2. Contains YYYYMMM-MMM, eg. 2013Mar-Apr
#Format3. Regular format eg. 2012Jul


#Extracting Days for pubDate that has date entries.
```{r}
#We shall ignore index number 12 since it's a year data entry
index_of_format1 <- grep("[0-9]+$", pubDate)
index_of_format1
value_of_format1 <- pubDate[index_of_format1]
value_of_format1 <- substring(value_of_format1, 8)
value_of_format1
```



#Taking only the latter month abbreviation for datas that have YYYYMMM-MMM formats
```{r}
index_of_format2 <- which(nchar(pubDate)==11)
index_of_format2
value_of_format2 <- pubDate[index_of_format2]
trimmed_value_of_format2 <- substr(value_of_format2, 9,11)
trimmed_value_of_format2
length(trimmed_value_of_format2)
```



```{r}
library(lubridate)
Year <- as.integer(substr(pubDate, 1,4))
Month <- substr(pubDate_without_space, 5,7)
#Replace values for special index 2,4,13,18
Month[2] <- trimmed_value_of_format2[1]
Month[4] <- trimmed_value_of_format2[2]
Month[13] <- trimmed_value_of_format2[3]
Month[18] <- trimmed_value_of_format2[4]

#Convert Month Names to Month in Numeric Forms
nMonth <- as.integer(match(Month,month.abb))

#Creating Pub_ID
Pub_ID <- seq(from = 1, to = length(pubDate)) + 100
Pub_ID <- paste("Pub_ID", Pub_ID, sep = '')

pubDate.df <- data.frame(Pub_ID=numeric(),
                         Journal_ID=numeric(),
                                  Date_Published=character(),
                                  Year_Published=integer(),
                                  Month_Published=integer(),
                                  Day_Published=integer(),
                                  stringsAsFactors = F)
for (i in 1:19) {
pubDate.df [i,1] <- Pub_ID[i]
pubDate.df [i,2] <- jID[i]
pubDate.df [i,3] <- pubDate[i]
pubDate.df[i,4] <- Year[i]
pubDate.df[i,5] <- nMonth[i]
pubDate.df[i,6] <- 31
}
pubDate.df

#Preparing Days values
pubDate.df$Day_Published <- days_in_month(nMonth)
pubDate.df$Day_Published[15] <- value_of_format1[2]
pubDate.df$Day_Published [16] <- value_of_format1[3]
pubDate.df$Day_Published [19] <- value_of_format1[4]
pubDate.df$Day_Published <- as.integer(pubDate.df$Day_Published)


#Replacing missing values in 12th row with December and 31 days
pubDate.df$Month_Published[12] <- as.integer(12)
pubDate.df$Day_Published[12] <- 31
pubDate.df


#Creating date in formats
pubDate.df$Date_Published_in_Format <- paste(pubDate.df$Year_Published, 
                                             pubDate.df$Month_Published,
                                             pubDate.df$Day_Published,
                                             sep="-")

pubDate.df$Date_Published_in_Format <- as.Date(pubDate.df$Date_Published_in_Format, format="%Y-%m-%d")
pubDate.df <- subset(pubDate.df, select = c(1,2,7,3,4,5,6))
pubDate.df
```

#Creating createdDate.df
```{r}
createdDate.df <- subset(article_with_DaysInfo.df, select = c(7, 2, 4,5,6))
colnames(createdDate.df) <- c("Date_Created_in_Format", "Journal_ID", "Date_Year", "Date_Month", "Date_Day")
createdDate.df$Created_ID <- seq(from = 1, to = nrow(createdDate.df)) + 100
createdDate.df$Created_ID <- paste("C_ID", createdDate.df$Created_ID, sep = '')

#Converting Formats
createdDate.df$Date_Year <- as.integer(createdDate.df$Date_Year)
createdDate.df$Date_Month <- as.integer(createdDate.df$Date_Month)
createdDate.df$Date_Day <- as.integer(createdDate.df$Date_Day)

createdDate.df <- subset(createdDate.df, select = c(6,2,1,3,4,5))
createdDate.df
```

#Creating days_elapsed.df
```{r}
days_elapsed.df <- data.frame(ED_Key=character(),
                              Journal_ID=numeric(),
                              Date_Created=character(),
                              Date_Published=character(),
                              Days_Elapsed=integer(),
                              stringsAsFactors = F)

dCreated <- article_with_DaysInfo.df$Date_Created
dPublished <- pubDate.df$Date_Published_in_Format
nElapsed <- difftime(dPublished, dCreated)
edKey <- seq(from = 1, to = length(dCreated)) + 100
edKey <- paste("ED", edKey, sep = '')

for (i in 1:length(dCreated)) {
  days_elapsed.df[i,1] <- edKey[i]
  days_elapsed.df[i,2] <- jID[i]
  days_elapsed.df[i,3] <- format(dCreated[i], "%Y-%m-%d")
  days_elapsed.df[i,4] <- format(dPublished[i], "%Y-%m-%d")
  days_elapsed.df[i,5] <- nElapsed[i]
}
#Converting data into correct data types
days_elapsed.df$Journal_ID <- as.integer(days_elapsed.df$Journal_ID)
days_elapsed.df$Date_Created <- as.Date(days_elapsed.df$Date_Created)
days_elapsed.df$Date_Published <- as.Date(days_elapsed.df$Date_Published)
days_elapsed.df
```
#Index 1 value turns to be negative, need to find a legit value to replace 
#Date_Created with PubMedPubDate PubStatus="received" in History Node 1

```{r}
Replaced_Year_xpath <- "//PubmedArticle/PubmedData/History/PubMedPubDate[@PubStatus='received']/Year"
Replaced_Year <- xpathSApply(xmlDOM,Replaced_Year_xpath, xmlValue)
#Replaced_Year

Replaced_Month_xpath <- "//PubmedArticle/PubmedData/History/PubMedPubDate[@PubStatus='received']/Month"
Replaced_Month <- xpathSApply(xmlDOM,Replaced_Month_xpath, xmlValue)
#Replaced_Month

Replaced_Day_xpath <- "//PubmedArticle/PubmedData/History/PubMedPubDate[@PubStatus='received']/Day"
Replaced_Day <- xpathSApply(xmlDOM,Replaced_Day_xpath, xmlValue)
#Replaced_Day

Replaced_Date <- paste(Replaced_Year[1],
                       Replaced_Month[1],
                       Replaced_Day[1], 
                         sep="-")
days_elapsed.df$Date_Created[1] <- Replaced_Date
days_elapsed.df$Days_Elapsed[1] <- difftime(days_elapsed.df$Date_Published[1], days_elapsed.df$Date_Created[1])

days_elapsed.df$Date_Created[1] <- as.Date(days_elapsed.df$Date_Created[1], format="%Y-%m-%d")

#Adding the leading zeros to the Journal_ID 
days_elapsed.df$Journal_ID[2] <- paste(0, days_elapsed.df$Journal_ID[2], sep = '')
days_elapsed.df$Journal_ID[3] <- paste(0, days_elapsed.df$Journal_ID[3], sep = '')
days_elapsed.df$Journal_ID[6] <- paste(0, days_elapsed.df$Journal_ID[6], sep = '')
days_elapsed.df$Journal_ID[9] <- paste(0, days_elapsed.df$Journal_ID[9], sep = '')
days_elapsed.df$Journal_ID[10] <- paste(0, days_elapsed.df$Journal_ID[10], sep = '')
days_elapsed.df$Journal_ID[16] <- paste(0, days_elapsed.df$Journal_ID[16], sep = '')
days_elapsed.df$Journal_ID[6] <- paste(0, days_elapsed.df$Journal_ID[6], sep = '')
days_elapsed.df

```

```{r}
library(dplyr)
library(lubridate)
library(zoo)
#days_elapsed.df[order(days_elapsed.df$Date_Published),]

#days_elapsed.df <- days_elapsed.df %>% group_by(Month_Published=floor_date(days_elapsed.df$Date_Published, "month"))
#days_elapsed.df
days_elapsed.df$Month_Published <- as.integer(substr(days_elapsed.df$Date_Published, 6,7))
days_elapsed.df$Quarter_Published <- quarters(days_elapsed.df$Date_Published)
days_elapsed.df$Year_Published <- as.integer(substr(days_elapsed.df$Date_Published, 1,4))
#days_elapsed.df <- group_by(days_elapsed.df$Month_Published)
days_elapsed.df
```
#Creating dataframes for Yearly and Quarterly Average Elapsed Days

```{r}
Yearly_ED.df <- aggregate(days_elapsed.df$Days_Elapsed, list(days_elapsed.df$Year_Published), FUN=mean)
colnames(Yearly_ED.df) <- c("Year", "Y_AVG_ED")
Yearly_ED.df
Quarterly_ED.df <- aggregate(days_elapsed.df$Days_Elapsed, list(days_elapsed.df$Quarter_Published), FUN=mean)
colnames(Quarterly_ED.df) <- c("Quarter", "Q_AVG_ED")
Quarterly_ED.df
```
#Merging days_elapsed.df with numArti.df, Yearly_ED.df, and Yearly_ED.df to 
#form the journal_fact.df
```{r}
journal_fact.df <- days_elapsed.df
journal_fact.df$Pub_ID <- pubDate.df$Pub_ID
journal_fact.df$Created_ID <- createdDate.df$Created_ID
journal_fact.df$ED_Key <- days_elapsed.df$ED_Key
```


```{r}
journal_fact.df <- merge(x=journal_fact.df, y=Yearly_ED.df,
                        by.x="Year_Published",by.y="Year",
                        all.y=TRUE)
journal_fact.df <- merge(x=journal_fact.df, y=Quarterly_ED.df,
                        by.x="Quarter_Published",by.y="Quarter",
                        all.y=TRUE)
journal_fact.df <- merge(x=journal_fact.df, y=numArti.df,
                       by.x="Journal_ID",
                      all.x=TRUE)
#journal_fact.df <- merge(x=journal_fact.df, y=pubDate.df,
#                        by.x="Journal_ID",by.y="Journal_ID",
#                       all.y=TRUE)


journal_fact.df <- subset(journal_fact.df, select=c(1,9,10,4,5,6,2,7,12,11,13))

#Change date type to character so journal_fact.df is ready to be 
#loaded into Star Schema with date in readable format.
journal_fact.df$Date_Created <- as.character(journal_fact.df$Date_Created)
journal_fact.df$Date_Published <- as.character(journal_fact.df$Date_Published)
journal_fact.df
```
#Change a few dates into character types so the tables in the star schema has 
#readable dates values instead of 5-digit values.
```{r}
#for Journal_fact, journal_fact.df is ready

#for Days_Elapsed.dim
days_elapsed_for_Star.df <- days_elapsed.df
days_elapsed_for_Star.df$Date_Created <- as.character(days_elapsed_for_Star.df$Date_Created)
days_elapsed_for_Star.df$Date_Published <- as.character(days_elapsed_for_Star.df$Date_Published)

#for Created_Date.dim
createdDate_for_Star.df <- createdDate.df
createdDate_for_Star.df$Date_Created_in_Format <-as.character(createdDate_for_Star.df$Date_Created_in_Format)

#Adding the leading zeros to the Journal_ID 
createdDate_for_Star.df$Journal_ID[2] <- paste(0, createdDate_for_Star.df$Journal_ID[2], sep = '')
createdDate_for_Star.df$Journal_ID[3] <- paste(0, createdDate_for_Star.df$Journal_ID[3], sep = '')
createdDate_for_Star.df$Journal_ID[6] <- paste(0, createdDate_for_Star.df$Journal_ID[6], sep = '')
createdDate_for_Star.df$Journal_ID[9] <- paste(0, createdDate_for_Star.df$Journal_ID[9], sep = '')
createdDate_for_Star.df$Journal_ID[10] <- paste(0, createdDate_for_Star.df$Journal_ID[10], sep = '')
createdDate_for_Star.df$Journal_ID[16] <- paste(0, createdDate_for_Star.df$Journal_ID[16], sep = '')
createdDate_for_Star.df$Journal_ID[6] <- paste(0, createdDate_for_Star.df$Journal_ID[6], sep = '')
createdDate_for_Star.df

#for Publication_Date.dim
pubDate_for_Star.df <- pubDate.df
pubDate_for_Star.df$Date_Published_in_Format <-as.character(pubDate_for_Star.df$Date_Published_in_Format)

```


#Populating the dimensional model for journal.db
```{r}
#Write Journal_fact
dbWriteTable(jcon, "Journal_fact",journal_fact.df, overwrite = T, row.names =F)

#Write Number_Article_dim
dbWriteTable(jcon, "Number_Article_dim", numArti.df, overwrite = T, row.names =F)

#Write Publication_Date_dim
dbWriteTable(jcon, "Publication_Date_dim", pubDate_for_Star.df, overwrite = T, row.names =F)

#Write Created_Date_dim
dbWriteTable(jcon, "Created_Date_dim", createdDate_for_Star.df, overwrite = T, row.names =F)

#Write Days_Elapsed_dim
dbWriteTable(jcon, "Days_Elapsed_dim", days_elapsed_for_Star.df, overwrite = T, row.names =F)
```

#Show first 10 rows of the Journal_fact Table
```{sql connection=jcon}
select * from Journal_fact
```
#Show first 10 rows of the Number_Article_dim Table
```{sql connection=jcon}
select * from Number_Article_dim limit 10
```
#Show first 10 rows of the Publication_Date_dim Table
```{sql connection=jcon}
select * from Publication_Date_dim limit 10
```

#Show first 10 rows of the Created_Date_dim Table
```{sql connection=jcon}
select * from Created_Date_dim limit 10
```

#Show first 10 rows of the Days_Elapsed_dim Table
```{sql connection=jcon}
select * from Days_Elapsed_dim limit 10
```
#Part3
#Adjusting the Journal_fact table
#Reason: 2 obvious outliers found in the Days_Elapsed column which should be 
#avoided in order to explore the pattern without getting biased.

#First: Adjust the journal_fact.df first 
```{r}
journal_fact.df_without_outliers <- subset(journal_fact.df, Days_Elapsed != 1 & 
                                   Days_Elapsed != 355)
journal_fact.df_without_outliers
```
#Second: adjusting average days_elapsed for values 1 and 355 in Q1 and Q3
```{r}
new_Quarterly_ED.df <- aggregate(journal_fact.df_without_outliers$Days_Elapsed, 
                                 list(journal_fact.df_without_outliers$Quarter_Published), 
                                 FUN=mean)
colnames(new_Quarterly_ED.df) <- c("Quarter", "Q_AVG_ED")
new_Quarterly_ED.df
```



#Delete the rows in Journal_fact table that has Days_Elapsed values equivalent
#to 1 and 355
```{sql connection=jcon}
DELETE FROM Journal_fact
WHERE
Days_Elapsed=1 
OR
Days_Elapsed=355;
```


#Adjust Average values for 1 and 355's matching quarters Q1 and Q4 
#in the Journal_fact table with Q1=36.00, Q4=52.75 in days_elapsed
```{sql connection=jcon}
UPDATE Journal_fact
SET Q_AVG_ED=36.00 WHERE Quarter_Published="Q1"
```

#SET Q_AVG_ED=52.75 WHERE Quarter_Published="Q4"
```{sql connection=jcon}
UPDATE Journal_fact
SET Q_AVG_ED=52.75 WHERE Quarter_Published="Q4"
```


```{sql connection=jcon}
SELECT * FROM Journal_fact
```


#Exploring Pattern
```{sql connection=jcon}
SELECT Quarter_Published, count(Pub_ID) AS Numbers_of_Publications FROM Journal_fact
GROUP BY Quarter_Published
```
#Line graph that demonstrates a pattern
#Number of publications incerases in Q2 and Q3 during summer and fall season
#whereas
#Number of publication decreases from Q4 to Q1 in winter time and early spring

#I would expect that a better pattern could be found if the publication date 
#and the created date of the journal are defined more specifically and more clearly 
#at the time when the date data was collected in the first place
```{r}
library(ggplot2)
new_Quarterly_ED.df$Q_AVG_ED <- round(new_Quarterly_ED.df$Q_AVG_ED, 2)

line_chart <- ggplot(new_Quarterly_ED.df, aes(x=`Quarter`, y=`Q_AVG_ED`, group = 1)) +
geom_line() +
labs(x = "Quarter_Published", y = "Average_Days_Elapsed") +
ggtitle("Number of Publications by Quarter") +
geom_text(aes(label=`Q_AVG_ED`, vjust= -0.5))
line_chart
```


#Showing the original line-graph that actually looks like a pattern is shown 
#however it's the one that I felt not comfortale drawing conclusion with
#due to the outliers
```{r}
library(ggplot2)
journal_fact.df$Q_AVG_ED <- round(journal_fact.df$Q_AVG_ED, 2)

line_chart <- ggplot(journal_fact.df, aes(x=`Quarter_Published`, y=`Q_AVG_ED`, group = 1)) +
geom_line() +
labs(x = "Quarter_Published", y = "Average_Days_Elapsed") +
ggtitle("Number of Publications by Quarter") +
geom_text(aes(label=`Q_AVG_ED`, vjust= -0.5))
line_chart
```



```{r}
dbDisconnect(jcon)
```
